Supplementary Material - Using high-resolution variant frequencies to empower clinical genome interpretation

### 2016-11-01 manuscript/SupplementaryMaterial.Rmd
Removed more words
More reformatting
Edits from JW

where *maximum genetic contribution* represents the proportion of all cases that are attributable to the gene under evaluation, and *maximum allelic contribution* represents the proportion of cases attributable to that gene that are attributable to an individual variant. The maximum frequency of a recessive disease causing variant in the population was therefore defined as:  
*max credible AF* = $\sqrt(prevalence) \times$ *maximum allelic contribution* $\times \sqrt(maximum genetic contribution) \times 1/\sqrt(penetrance)$

@@

It is often difficult to obtain accurate penetrance information for reported variants, and it is also difficult to know what degree of penetrance to expect or assume in newly discovered pathogenic variants. In this work we uniformly apply a value of 50% penetrance (i.e. assuming variant penetrance is no lower than 50%) equivalent to that reported for our HCM example variant, and the lowest we found reported for any of our examples or reported accross our studied disorders. We recognise several other approaches that can be used to deal with the issues of penetrance, these include: setting a penetrance level equivalent to the minimum that is 'clinically actionable' for a disorder; lowering the penetrance if reduced penetrance is expected in a family; or using a two tiered approach, initially searching for a high-moderate penetrance variant but allowing for a lower-penetrance variant in a second pass. We believe that the ease of re-calculating our "maximum credible population allele frequency" lends itself to any of these approaches. We provide an online calculator to facilitate the exploration of these parameters (<https://jamesware.shinyapps.io/alleleFrequencyApp/>). If there are large case and control populations for a disease and the diease prevalence is known, we can use these to estimate penetrance `r citep("10.1126/scitranslmed.aad5169")`.

>>>>

where *maximum genetic contribution* represents the proportion of all cases that are attributable to the gene under evaluation, and *maximum allelic contribution* represents the proportion of cases attributable to that gene that are attributable to an individual variant. This can be rearranged to give the maximum frequency of a recessive disease causing variant in the population as:
*max credible AF* = $\sqrt($*prevalence*$) \times$ *maximum allelic contribution* $\times \sqrt($*maximum genetic contribution*$) \times 1/\sqrt($*penetrance*$)$

@@

It is often difficult to obtain accurate penetrance information for reported variants, and it is also difficult to know what degree of penetrance to expect or assume in newly discovered pathogenic variants. In this work we uniformly apply a value of 50% penetrance for inherited cardiac conditions (i.e. assuming variant penetrance is no lower than 50%) equivalent to that reported for our HCM example variant, and the lowest we found reported for any of our examples or reported accross our studied disorders. We recognise several other approaches that can be used to deal with the issues of penetrance, these include: setting a penetrance level equivalent to the minimum that is 'clinically actionable' for a disorder; lowering the penetrance if reduced penetrance is expected in a family; or using a two tiered approach, initially searching for a high-moderate penetrance variant but allowing for a lower-penetrance variant in a second pass. We believe that the ease of re-calculating our "maximum credible population allele frequency" lends itself to any of these approaches. We provide an online calculator to facilitate the exploration of these parameters (<https://jamesware.shinyapps.io/alleleFrequencyApp/>). If there are large case and control populations for a disease and the diease prevalence is known, we can use these to estimate penetrance `r citep("10.1126/scitranslmed.aad5169")`.
